---
title: "Problem Set 3"
author: "Thomas O'Leary and Amanda Jones"
date: "9/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Structured population models

```{r comadre, echo = TRUE, eval = TRUE}
load("COMADRE_v.3.0.0.RData")
# save stage transition matrix for the whales
whale_mat <- comadre$mat[2136][[1]]$matA
```

### Sperm whales

We are looking _Physeter macrocephalus occupying_, specifically a population of sperm whales in the Caribbean, to attempt to determine whether or not that population is expected to increase in size over time. We want to look at the the proportion between age classes, in this case, unweaned calves and adults (excluding mature males). We are also curious how small changes in birth, and survival rates will impact growth trends. We would like to know if the population is healthy/stable and if it is not, which life stage will have the greatest impact for conservation strategies. 

### Eigenvalue

```{r eigenvalue, echo = FALSE, eval = TRUE}
eigenvalue <- eigen(whale_mat)
dom_eigen <- max(eigenvalue$values)
```

```{r eigenvalue, echo = TRUE, eval = TRUE}
# dominant eigenvalue
dom_eigen
```

### Stable age distribution (i.e. eigenvector)

```{r eigenvector, echo = FALSE, eval = TRUE}
# get the eigenvector associated with dominant eigenvalue
# first value refers to the calves and second value refers to the adults
eigenvector <- eigenvalue$vectors[, 1]

# calculate the stable age distribution of calves to adults
stable_age_ratio <- eigenvector[2]/eigenvector[1]

prop_calves <- round(eigenvector[1]/(eigenvector[1] + eigenvector[2]), 
                     digits = 3)
prop_adults <- round(eigenvector[2]/(eigenvector[1] + eigenvector[2]), 
                     digits = 3)
```
There is a stable age distribution of approximately 79% adults and 21% unweaned calves.

```{r, echo = TRUE, eval = TRUE}
stable_age_ratio
prop_calves
prop_adults
```


### Sensitivity

```{r}
#install.packages("popbio")
library(popbio)

sensitivity(whale_mat)
```



### Elasticity

```{r}
elasticity(whale_mat)
```

### What does the eigenvalue, the eigenvector, sensitivity, and elasticity tell you about the population?

The eigenvalue is 0.97 which means that the population size is declining because it is less than one. The eigenvector tells us the stable age dsitribution. There are 3.7 times as many adults as there are unweened calves, or there are aprroximately 79% adults and 21% unweened calves. The sensitivity tells us how much a small perturbation in each parameter will affect the population growth rate overall. So the highest sensitivity value is the adults that remain adults (_i.e._ continue to survive). The reason that this might be the highest sensitivity value is that the whales are "suddenly immortal" so that would clearly have a profound effect on the population size. But it is not biologically realistic. Although the survival of adults is still probably an important parameter. 

The elasticity is the proportional sensitivity. 

So  


### Address question proposed using eigenvalue, eigenvector, elastisity, & sensitivity

